---
title: "Project proposal"
author: "Team name"
output: pdf_document
---

```{r load-packages, message = FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library("ggpubr")
my_data <- read.csv(" ", header=TRUE, stringsAsFactors=FALSE)
head(my_data, " ")
```

```{r setup, include = FALSE}
opts_chunk$set(echo=FALSE) # hide source code in the document
library("ggpubr")
ggqqplot(my_data$time, ylab = "Time")
ggqqplot(my_data$amplitude, ylab = "Electrode")
library("ggpubr")
ggscatter(my_data, x = "Time", y = "Electrode", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Time", ylab = "Electrode")
```

## 1. Introduction
The Differences Between Various Anti-____ on Neurons using Multielectrode Array.
Memory is the faculty of the mind by which data or information is encoded, stored, and retrieved when needed. A memory is formed when neurons transfer electronic signals between one to another. It may be caused by a new experience or reliving previous ones, which occur by transferring electric pulses between neurons that were originally involved in forming the memory. 

Throughout history, humans has always tried to improve that ability.
The presented research's aim is to find a drug that is proved to improve memory. 
Our specific problem is the last step to do so: With two specific drugs (AP5, FGL) we would like to determine whether they are efficient or not, and if so, which one is more efficient?

Our approach is to find the correlation between different electrodes on a single petri dish and later to compare between the different perti dish.
Every petri dish is given a different treatment with different dosage. A few of them are the control group and are not given a treatment at all.

## 2. Data


## 3. Preliminary results


## 4. Data analysis plan
Our plan: 
1. Clean up the data: The data is automatically recorded and due to some technical failures has some NULL rows that might interrupt the analysis. 

2. Combine the different data files.

3. Find correlation between electrodes per petri dish. This process has to be done several times, for each step of the experiment: before treatment, during treatment and during the night after the treatment. 
This will help us determine which neurons formed connection before and during treatment and if so, has it improved?

4. Find correlation between different petri dishes. This will help us determine if the treatment is efficient, and which one is the most. 

We will analyse the connections formed between two electrode in a petri dish. Electrodes that are forming more electrical signals between them, have a stronger connection. 
After that we will analyse the correlation between different petri dishes.

For this analysis, we will create a correlation matrix per petri, probably by Pearson's Correlation. 

The correlation matrix will help us determine if the connections formed became stronger with treatment, and if so, we will be able to say the treatment help improving memory. 

The data is divided to 3 time periods:
1. Before treatment
2. During treatment
3. After treatment (at night)

Each group member will analyse one of these time periods. In the end, we will combine our conclusions in order to conclude the final conclution.

## Appendix

### Data README

```{r include_data_readme, comment=''}
cat(readLines('../data/README.md'), sep = '\n')
```

### Source code

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```
