---
title: |
  \textbf{The Differences Between Various Memory-Improvement Treatments on Neurons\\ using Multielectrode Array}
author: "Team P"
date: "19 June 2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
subtitle: Final Project Report
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages, message = FALSE, echo=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library(meaRtools)
```
The code repository can be found [here](https://github.com/omertol/R-project)

## Introduction

In general, memory is the process of encoding, storing, and retrieving experiences and knowledge.
We would like to generally describe the process of forming a memory.
This background will help us explain our project.
Please note that in the following document, 'neurons' will only refer to those found in the brain, and not to those found in the nervous system.

Different groups of neurons, responsible for different thoughts or perceptions, drift in and out of action.
A 'group' (called a synapse) is formed when two neurons create a connection by transferring electric signals between one another.
These connections can be made stronger or weaker depending on when and how often they have been activated in the past.
Active connections tend to get stronger, whereas inactive ones get weaker and can eventually disappear entirely.
Retrieving a memory is the reactivation of the specific synapse that was originally formed when the experience first happend.

Throughout history, humans have always tried to improve the ability to strengthen the connections between neurons, while also creating new ones.
The aim of the presented research is to find a drug that is proven to improve memory, meaning maintaining the current connections for a longer period of time, even when not used.
Neurons are bred in a well with electrodes in a laboratory.
Each well is given a different treatment at a different dosage.
A few of them are the control group and are not given a treatment at all.
The electrodes monitor and document every connection made in each of the wells.
When a neuron sends an electric signal to another neuron, an electric potential difference (voltage) is created and recorded by the electrodes.

The problem we are tackling is the last step of the process: we would like to determine whether the drugs (AP5, FGL) are efficient or not, and if so, which one is more efficient.
Finding an efficient memory-improvement treatment might be the right path to prevent Alzheimer's Disease and dementia.

Our approach is to first clean and combine the different datasets for each episode of time. Then, we will find the Pearson's correlation between different electrodes on a single well. Finally, we will compare the different wells using the ANOVA test and perform a post-hoc test in the Tukey method.

## Data Overview

We have extended the amount of data we used for this project since the proposal was submitted and it now includes 9 data files (.csv), produced automatically by the multielectrode array.
Each file was recorded at different episodes of the experiment: before treatment (1 file, starts with bl); during treatment (5 files, starts with ac); after treatment (during the night) (3 files, starts with on). The `README.md` was updated accordingly.

Each of them contains 3 main features:

-   Time (s) [Float]: Seconds passed from the start of the recording until a signal was sent.

-   Electrode [String]: The electrode on which the neuron that sent the signal is placed.
    \
    This column contains the well's ID and the electrode's ID, separated by a lower dash as the following format: [Well name]\_[Electrode name]

-   Amplitude (mV) [Float]: The electronic signals strength (voltage).

There is a total of 24 wells, each containing 16 electrodes.
Neurons activity was recorded at selected times during a period of 24 hours, for 11 minutes each.

The data also contains some additional information about the different wells, the research, the equipment used, etc.
All of these are irrelevant to the problem we are tackling.
For additional information about the above, please refer to the `README.md` file in the `Data` folder in the project repository.

## Methods and Results

## Limitations and Future Work

Prior work indicates there is more than one neuron per electrode. Finding the correlation between electrodes helps us understand the connections between the synapses. It cannot, however, teach us about the connections within the synapse itself, that may become stronger due to the treatment.
Generalizing this solution beyond our sample will require a more accurate way to map the connections between neurons, so that we could further understand the treatments impact.

With additional time and resources, first we would have consulted with prof. Knafo on how to identify a single neuron in a synapse and try to find the connections between the neurons themselves, rather than between synapses. Second, we would have tried to find and implement a more accurate and efficient method to evaluate the connections between the neurons. Finally, we would have extended the duration of the recording for more accurate results. With these additional steps, we would have been able to conclude the treatments' efficiency better, not only during the 24 hours following the treatment, but in the long-term as well.

```{=tex}
\begin{center}
Special thanks to prof. Shira Knafo, who kindly agreed to share the data from her research: "Understanding Neural Development with Multielectrode Array" with us. The data was imperative to this project, and proved to be invaluable during its making.
\end{center}
```
