---
	title: |
	  \textbf{The Differences Between Various Memory-Improvement Treatments on Neurons\\ using Multielectrode Array}
	author: "Team P"
	date: "19 June 2022"
	output:
	  word_document: default
	  html_document: default
	  pdf_document: default
	subtitle: Final Project Report
	editor_options:
	  markdown:
	    wrap: sentence
	---
	```{r setup, include=FALSE}
	knitr::opts_chunk$set(echo = FALSE)
	```
	```{r load-packages, message = FALSE, echo=FALSE}
	library(knitr)
	library(tidyverse)
	library(broom)
	library(htmltools)
	```
	```{=tex}
	\begin{center}
	The code repository can be found \href{https://github.com/omertol/R-project}{here}.
	\end{center}
	```
	## Introduction
  
	In general, memory is the process of encoding, storing, and retrieving experiences and knowledge.
	We would like to generally describe the process of forming a memory.
	This background will help us explain our project.
	Please note that in the following document, 'neurons' will only refer to those found in the brain, and not to those found in the nervous system.
	A 'group' (called a synapse) is formed when two neurons create a connection by transferring electric signals between one another.
	These connections can be made stronger or weaker depending on when and how often they have been activated in the past.
	Active connections tend to get stronger, whereas inactive ones get weaker and can eventually disappear entirely.
	
	The aim of prof. Knafo research is to find a drug that is proven to improve memory, meaning maintaining the current connections for a longer period of time, even when not used.
	Neurons are bred in a well with electrodes in a laboratory. Each well is given a different treatment at a different dosage.A few of them are the control group and are not given a treatment at all.
	The electrodes monitor and document every connection made in each of the wells.
	When a neuron sends an electric signal to another neuron, an electric potential difference (voltage) is created and recorded by the electrodes.
	
	The problem we are tackling is the last step of the process: we would like to determine whether the drugs (AP5, FGL) are efficient or not, and if so, which one is more efficient.
	Finding an efficient memory-improvement treatment might be the right path to prevent Alzheimer's Disease and dementia.
	
	Our approach is to first clean and combine the different datasets for each episode of time. Then, we will find the Pearson's correlation between different electrodes on a single well. Finally, we will compare the different wells using ANOVA tests and perform a post-hoc test in the Tukey method.
	
	## Data Overview
	
	We have extended the amount of data we used for this project since the proposal was submitted and it now includes 9 data files (.csv), produced automatically by the multielectrode array.
	Each file was recorded at different episodes of the experiment: before treatment (1 file, starts with bl); during treatment (5 files, starts with ac); after treatment (during the night) (3 files, starts with on). The `README.md` was updated accordingly.
	
	Each of them contains 3 main features:
	
	-   Time (s) [Float]: Seconds passed from the start of the recording until a signal was sent.
	
	-   Electrode [String]: The electrode on which the neuron that sent the signal is placed.
	    \
	    This column contains the well's ID and the electrode's ID, separated by a lower dash as the following format: [Well name]\_[Electrode name]
	
	-   Amplitude (mV) [Float]: The electronic signals strength (voltage).
	
	There is a total of 24 wells, each containing 16 electrodes.
	Neurons activity was recorded at selected times during a period of 24 hours, for 11 minutes each.
	
	The data also contains some additional information about the different wells,## the research, the equipment used, etc.
	All of these are irrelevant to the problem we are tackling.
	For additional information about the above,## please refer to the `README.md` file in the `Data` folder in the project repository.
	
	## Methods and Results
	
	At first, we attempted to find the answer to our research question using a KNN model. For each time period, we calculated the firing rate for each treatment by summing the firing rate of different wells who were given the same treatment. We calculated the change in the firing rate for adjacent time periods. Our hypotheses was that if there is a significant difference in the firing rate of each well with a specific treatment, the model will be able to determine which one is it. Therefore we could understand the treatment's influence over the neurons. 
	
	The model accuracy was low. The model could not determine which treatment was given to a well based on the change in its firing rate.
	We concluded the change in firing rate is not a way to measure the treatment's efficiency. An electrode might send signals consistently but with no response.
	
	We decided to try a different approach. We calculated every pair of electrodes' correlation within a single well using Pearson's correlation method. The correlation is stronger in electrodes that sent a signal during adjacent times. We repeated this process for every well at each time period. 
	
	To determine whether there is a significant difference between the treatments, and between a treatment to itself at different time periods, we decided to perform numerous ANOVA tests. For a p-value < 0.05, a null hypotheses is rejected in each of the tests. 
	
	The null hypotheses we wish to test were:
	1. There is no difference in correlation for any treatment.
	2. There is no difference in correlation for any time period.
	3. The interaction between treatment and time period does not effect the correlation.
	
	Because we would like to examine the relationship between the time and treatment, we compared two models:
	
	1. Additive two-way ANOVA: A model that does not predict any interaction between the time period and the treatment. 
	2. Two-way ANOVA with interaction: A model to test whether the interaction between treatment and time period influences the final correlation.
	
	To choose the best fitted model for our data between the two, we used AIC's method. The method showed that the second model is better fitted to our data. From interpreting its results, we could reject all three null hypotheses. 
	Yet, by diagnosing the regression's plots (fig.1), we can see that the residuals values are no randomized and the normal probability plot is not linear. Therefore the model is not well-fitted to the data. [fig.1]
	 The additional null hypotheses we wish to test were: 
	1. There is no difference in correlation for different distance values.
	2. The interaction between treatment and distance does not effect the correlation.
	3. The interaction between time period and distance does not effect the correlation.
	4. The interaction between time period, treatment and distance does not effect the correlation.
	
	The new three-way ANOVA model rejected only the two first hypotheses. Therefore we created another two-way ANOVA model, this time using the distance variable as a blocking variable. [fig.2]
	Lastly, we used the Tukey's Honestly-Significant-Difference (TukeyHSD) test to examine the difference between the treatments and between a treatment in its initial state to its end state. [tbl.1] , [tbl.2]
	
 By interpreting these tables we can see that:
	* There is a significant difference between the treatments and between each treatment to the control group
	* FGL exhibited a significant difference between its initial state to its end state.
	* AP5 did not exhibited a significant difference between its initial state to its end state.
	
	The following plots (fig.3) show the change in the correlation based on the distance of electrodes from one another at different time periods.
	[fig.3]
	This is a clear visualization to FGL's efficiency in comparison to AP5 and the control group. The correlation value increases after the FGL treatment, while the rest were in about the same level.
	In conclusion, FGL is proven to be the better memory-improving treatment of the two.
	

	## Limitations and Future Work
	Prior work indicates there is more than one neuron per electrode.
	Generalizing this solution beyond our sample will require a more accurate way to map the connections between neurons, so that we could further understand the treatments impact.
	
	With additional time and resources, first we would have consulted with prof. Knafo on how to identify a single neuron in a synapse and try to find the connections between the neurons themselves, rather than between synapses. Second, we would have tried to find and implement a more accurate and efficient method to evaluate the connections between the neurons. Finally, we would have extended the duration of the recording for more accurate results. With these additional steps, we would have been able to conclude the treatments' efficiency better, not only during the 24 hours following the treatment, but in the long-term as well.
	
	```{=tex}
	\begin{center}
	Special thanks to prof. Shira Knafo, who kindly agreed to share the data from her research: "Understanding Neural Development with Multielectrode Array" with us. The data was imperative to this project, and proved to be invaluable during its making.
	\end{center}
