---
title: |
  \textbf{The Differences Between Various Memory-Improvement Treatments on Neurons\\ using Multielectrode Array}
author: "Team P"
date: "19 June 2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
subtitle: Final Project Report
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages, message = FALSE, echo=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library(meaRtools)
```
The code repository can be found [here](https://github.com/omertol/R-project)

## Introduction

In general, memory is the process of encoding, storing, and retrieving experiences and knowledge.
We would like to generally describe the process of forming a memory.
This background will help us explain our project.
Please note that in the following document, 'neurons' will only refer to those found in the brain, and not to those found in the nervous system.

Different groups of neurons, responsible for different thoughts or perceptions, drift in and out of action.
A 'group' (called a synapse) is formed when two neurons create a connection by transferring electric signals between one another.
These connections can be made stronger or weaker depending on when and how often they have been activated in the past.
Active connections tend to get stronger, whereas inactive ones get weaker and can eventually disappear entirely.
Retrieving a memory is the reactivation of the specific synapse that was originally formed when the experience first happend.

Throughout history, humans have always tried to improve the ability to strengthen the connections between neurons, while also creating new ones.
The aim of the presented research is to find a drug that is proven to improve memory, meaning maintaining the current connections for a longer period of time, even when not used.
Neurons are bred on a well with electrodes in a laboratory.
Every well is given a different treatment at a different dosage.
A few of them are the control group and are not given a treatment at all.
The electrodes monitor and document every connection made in each of the wells.
When a neuron send electric signal to another, an electric potential difference (voltage) is created and recorded by the electrodes.

Our specific problem is the last step: we would like to determine whether the drugs (AP5, FGL) are efficient or not, and if so, which one is more efficient.
Finding an efficient memory-improvement treatment, might be the right path to prevent Alzheimer and dementia.

Our approach is to combine the different dataset for each episode of time, find the Pearson's correlation between different electrodes on a single well and to later compare between the different wells.

## Data Overview

We have 6 data files (.csv), produced automatically by the multielectrode array.
Each recorded at different episodes of the experiment: before treatment (1 file, starts with bl), during treatment (2 files, starts with ac) and after treatment (during the night) (3 files, starts with on).

Each of them contains 3 main features:

-   Time (s) [Float]: Seconds passed from the start of the recording until a signal was sent.

-   Electrode [String]: The electrode on which the neuron that sent the signal is placed.
    \
    This column contains the well's ID and the electrode's ID, separated by a lower dash as the following format: [Well name]\_[Electrode name]

-   Amplitude (mV) [Float]: The electronic signals strength (voltage).

Total of 6 wells, each contains 16 electrodes.
Neurons activity was recorded at selected times during 24 hours for one hour each.

The data also contains some additional information about the different wells, the research, the equipment used, etc.
All of these are irrelevant to the problem were tackling.

## Methods and Results

## Limitations and Future Work

Due to limited computing resources, we couldn't efficiently compute the correlations between the electrodes in the most accurate way.
The data was too large for our personal computers and the other methods we found could not be applied to it due to its structure.
Creating a dataset whereas each electrode has its own column generated a dataset bigger to the power of sixteen from its original size.
Real-life data is obviously even bigger.
Generalizing this solution beyond our sample, might need enormous computing resources.

If we had more time, we would most definitely try to find a more accurate and efficient way to evaluate the connections between the synapses and implement it on the data using a more powerful resources.
In addition, if we could use a better equipment, we would have extended the data files and duration of the recording for more accurate results, so we could conclude the treatments efficiency in the long run, and not only the 24 hours following the treatment.
Moreover, prior work indicates there are more than one neuron on each electrode.
With addition time, we would have consulted with prof. Knafo on how to identify a single neuron in a synapse, and try to find the connection between the neurons themselves, rather than between a whole synapse to another.

```{=tex}
\begin{center}
Prof. Shira Knafo kindly agreed to share her research's "Understanding Neural Development with Multielectrode Array" data with us, which was used for this project.
\end{center}
```
