---
title: "Project proposal"
author: Team P
output: pdf_document
---

```{r load-packages, message = FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
```

```{r setup, include = FALSE}
opts_chunk$set(echo=FALSE) # hide source code in the document
```

## 1. Introduction
The Differences Between Various Memory-Improvement Treatments on Neurons using Multielectrode Arrays.

In general, memory is the process of encoding, storing, and retrieving experiences and knowledge.
We would like to generally describe the process of forming a memory. This background will help us explain our project.
Please note that in the following document, 'neurons' will only refer to those found in the brain, and not to those found in the nervous system.

The brain simmers with activity. Different groups of neurons, responsible for different thoughts or perceptions, drift in and out of action.
A 'group' is formed when two neurons create a connection - called synapses - by transfering electronic signals between one another.
These connections can be made stronger or weaker depending on when and how often they have been activated in the past. 
Active connections tend to get stronger, whereas those that aren’t used get weaker and can eventually disappear entirely.
Retrieving a memory is the reactivation of a specific group of neurons, formed from persistent changes in the strength of connections between neurons.

Throughout history, humans have always tried to improve that ability: strengthen the connections between neurons while creating new ones.
The presented research's aim is to find a drug that is proved to improve memory, meaning maintaining the current connections for a longer period of time, even when not used. 
Neurons are bred on a petri dish with electrodes in a laboratory. Every petri dish is given a different treatment with different dosage. A few of them are the control group and are not given a treatment at all.
The electrodes monitor and document every connection made in each of the petri dishes.
Our specific problem is the last step to do so: for two specific drugs (AP5, FGL) we would like to determine whether they are efficient or not, and if so, which one is more efficient.

Our approach is to find the correlation between different electrodes on a single petri dish and later to compare between the different petri dishes.


## 2. Data
The data includes six Excel files , each of them describes different time periods on the research.
The different time periods:
1. Before treatment - 1 Excel file ; area code "bi"
2. During treatment - 2 Excel files ; area code "ac"
3. After treatment (at night) - 3 Excel files ; area code "on"
Every Excel file contains records of the time and the Petri dish number_Electrode number who transfering electronic signals, also called "potential action".



## 3. Preliminary results


## 4. Data analysis plan
Our plan: 
1. Clean up the data: The data is automatically recorded and due to some technical failures has some NULL rows that might interrupt the analysis. 

2. Combine the different data files.

3. Find correlation between electrodes per petri dish. This process has to be done several times, for each step of the experiment: before treatment, during treatment and during the night after the treatment. 
This will help us determine which neurons formed connection before and during treatment, and if so, has it gone stronger or weaker?

4. Compare different petri dishes. This will help us determine if the treatment is efficient, and which one is the most. 

We will analyze the connections formed between two electrodes in a petri dish. Electrodes that are recording frequent electric signals between one another, have a stronger connection. 
After that we will analyze the correlation between different petri dishes.
For this analysis, we will create a correlation matrix per petri, probably by Pearson's Correlation. We still need to determine which method to use to compare petri dished. 

The correlation matrix will help us determine if the connections formed became stronger with treatment, and if so, we will be able to say the treatment help improve memory. 
## הפסקה הבאה נלקחה לחלק 2- לדעתי אפשר למחוק (סתיו)
The data is divided to 3 time periods:
1. Before treatment
2. During treatment
3. After treatment (at night)

Each group member will analyze one of the time periods. In the end, we will combine our conclusions in order to find the best treatment from the given ones.

## Appendix

### Data README

```{r include_data_readme, comment=''}
cat(readLines('../data/README.md'), sep = '\n')
```

### Source code

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```
