---
title: "Project proposal"
author: Team P
output: pdf_document
---

```{r load-packages, message = FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
```

```{r setup, include = FALSE}
opts_chunk$set(echo=FALSE) # hide source code in the document
```

## 1. Introduction
The Differences Between Various Memory-Improvement Treatments on Neurons using Multielectrode Arrays.

In general, memory is the process of encoding, storing, and retrieving experiences and knowledge.
We would like to generally describe the process of forming a memory. This background will help us explain our project.
Please note that in the following document, 'neurons' will only refer to those found in the brain, and not to those found in the nervous system.

The brain simmers with activity. Different groups of neurons, responsible for different thoughts or perceptions, drift in and out of action.
A 'group' (called a synapse) is formed when two neurons create a connection by transferring electric signals between one another.
These connections can be made stronger or weaker depending on when and how often they have been activated in the past. 
Active connections tend to get stronger, whereas inactive ones get weaker and can eventually disappear entirely.
Retrieving a memory is the reactivation of a specific group of neurons that was originally formed when the experience first happend.

Throughout history, humans have always tried to improve the ability to strengthen the connections between neurons, while also creating new ones.
The aim of the presented research is to find a drug that is proven to improve memory, meaning maintaining the current connections for a longer period of time, even when not used. 
Neurons are bred on a petri dish with electrodes in a laboratory. Every petri dish is given a different treatment at a different dosage. A few of them are the control group and are not given a treatment at all.
Neurons are bred on electrodes that monitor and document every connection made in each of the petri dishes. When a neuron send electric signal to another, an electric potential difference (voltage) is created and recorded by the electrodes.
Our specific problem is the last step: we would like to detemine whether the drugs (AP5, FGL) are efficient or not, and if so, which one is more efficient.

Our approach is to find the correlation between different electrodes on a single petri dish and to later compare between the different petri dishes.


## 2. Data
The data includes six Excel files , each of them describes different time periods on the research.
The different time periods:
1. Before treatment - 1 Excel file ; area code "bi"
2. During treatment - 2 Excel files ; area code "ac"
3. After treatment (at night) - 3 Excel files ; area code "on"
Every Excel file contains records of the time and the Petri dish number_Electrode number who transfering electronic signals, also called "potential action".



## 3. Preliminary results


## 4. Data analysis plan
Our plan: 
1. Clean up the data: The data is automatically recorded and due to some technical failures has some NULL rows that might interrupt the analysis. 

2. Combine the different data files.

3. Find correlation between electrodes per petri dish. This process must be done several times for each step of the experiment: before treatment, during treatment and during the night after the treatment. 
This will help us determine which neurons formed a connection before and during treatment, and if so, has it grown stronger or weaker?

4. Compare different petri dishes. This will help us determine which of the two treatments is most efficient, if at all.

We will analyze the connections formed between two groups of neurons, each lies on different electrode in a petri dish. Electrodes that recorde frequent electric signals between one another will be deemed to have had a stronger connection. 
After that we will analyze the correlation between different petri dishes.
For this analysis, we will create a correlation matrix per petri, probably according to Pearson's Correlation. We will determine which method to use to compare petri dishes in the future. 

The correlation matrix will help us determine if the connections formed became stronger following the treatment. If that is the case, we will be able to claim the treatment caused an improvement in memory. 

Each group member will analyze one of the time periods. In the end, we will combine our conclusions to find the best treatment from the given ones.

## Appendix

### Data README

```{r include_data_readme, comment=''}
cat(readLines('../data/README.md'), sep = '\n')
```

### Source code

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```
